---
title: "fact-check"
output: html_document
date: "2023-02-22"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
library(sf)
library(mapview)
library(tidycensus)
options(digits=3)
options(scipen=999)
'%notin%' <- Negate('%in%')


#does a standard group_by and count() with percentage
grouper <- function(input_df, group_by_column, new_column_name = "n()"){
  output_df <- input_df %>%
    group_by(.data[[group_by_column]]) %>%
    summarise(temp_count = n()) %>%
    mutate(percent = temp_count/sum(temp_count)*100) %>%
    arrange(desc(percent)) %>%
    rename(!!new_column_name := temp_count)
  return(output_df)
}

#group/counts every column in input dataframe
group_count <- function(input_df, group_column_name='n()', state_filter=NA, start_col = 1){
  column_names <- colnames(input_df)
  if(!is.na(state_filter)){
    input_df <- input_df %>%
      filter(state == state_filter)
  }
  for (column in column_names[start_col:length(column_names)]){
    output <- grouper(input_df, column, group_column_name)
    print(output)
  }
}

grouper_sum <- function(input_df, group_by_column, sum_column, new_column_name = "n()"){
  output_df <- input_df %>%
    group_by(.data[[group_by_column]]) %>%
    summarise(temp_count = sum(.data[[sum_column]])) %>%
    mutate(percent = temp_count/sum(temp_count)*100) %>%
    arrange(desc(percent)) %>%
    rename(!!new_column_name := temp_count)
  return(output_df)
  
}

#lowers case of every character column in a dataframe
lower_df <- function(input_df){
  names <- colnames(input_df)
  output_df <- input_df
  names <- colnames(output_df)
  for (name in names){
    if (is.character(output_df[[name]])){
      output_df[[name]] <- tolower(output_df[[name]])
      #print('yes')
    } else {
      output_df[[name]] <- output_df[[name]]
      #print('no')
    }
  }
  return(output_df)
}


round_df <- function(x, digits) {
    # round all numeric variables
    # x: data frame 
    # digits: number of digits to round
    numeric_columns <- sapply(x, mode) == 'numeric'
    x[numeric_columns] <-  round(x[numeric_columns], digits)
    x
}

```

## IMPORT PART 1 CRIME VICTIMS DATABASE

```{r}

victims <- read_csv('data/Part_1_Crime_Data_.csv') %>%
  #cleaning column names
  clean_names() %>%
  #lowering text in df
  lower_df()%>%
  #converting to lubridate
  mutate(crime_date_time = ymd_hms(crime_date_time)) %>%
  #making year month hour columns
  mutate(year = year(crime_date_time),
         month = month(crime_date_time),
         hour = hour(crime_date_time),
         date = as_date(crime_date_time, 'day')) %>%
  #removing lat/long because there are multiples and ethnicity because it is unreliable
  select(-x, -y, -row_id, -latitude, -longitude, -ethnicity) %>%#, -shape) %>%
  #filtering for year
  #filter(year >= 2011) %>%
  #splitting geo location
  separate(geo_location, c("latitude", "longitude"), ',') %>%
  #removing parenthesis
  mutate(latitude = gsub('\\(', '', latitude),
         longitude = gsub('\\)', '', longitude)) %>%
  #modifying ages that can't be or are very unlikely to be real
  mutate(age = case_when(
    age > 0 & age < 100 ~ age,
    TRUE ~ NA_real_
  )) %>%
  #classifying age ranges
  mutate(age_range = case_when(
    age > 0 & age <= 12 ~ "child",
    age <= 18 & age > 12 ~ "high school teen",
    age >= 18 & age <= 25 ~ "19 to 25",
    age >= 26 & age <= 34 ~ "26 to 34",
    age >= 35 & age <= 50 ~ "35 to 50",
    age > 50 & age < 75 ~ "51 to 74",
    age >= 75 & age < 85 ~ "75 to 84",
    age >= 85 ~ "85 to 100",
    TRUE ~ "NA"
  )) %>%
  #adding high_school_age classification
  mutate(high_school_age_range = case_when(
    age < 19 & age > 0 ~ "high school age",
    age >= 19 & age <= 25 ~ "19 to 25",
    age >= 26 & age <= 34 ~ "26 to 34",
    age >= 35 & age <= 50 ~ "35 to 50",
    age > 50 & age < 75 ~ "51 to 74",
    age >= 75 & age < 85 ~ "75 to 84",
    age >= 85 ~ "85 to 100",
    TRUE ~ "NA"
  )) %>%
  #classifying time of shooting
  mutate(time_range = case_when(
    hour <= 3 ~ "early morning",
    hour < 12 & hour > 3 ~ "morning",
    hour >= 12 & hour < 15 ~ "early afternoon",
    hour >= 13 & hour < 17 ~ "after school",
    hour >= 17 & hour < 20 ~ "evening",
    hour >= 20 ~ "night"
  ))  %>%
  #labeling quarter by month
    mutate(quarter = case_when(
    month %in% c(1,2,3) ~ 1,
    month %in% c(4,5,6) ~ 2,
    month %in% c(7,8,9) ~ 3,
    month %in% c(10,11,12) ~ 4
  )) %>%
  #labeling end date of each quarter for viz purposes
  mutate(quarter_date = case_when(
    quarter == 1 ~ ymd(paste0(year, "-3-31")),
    quarter == 2 ~ ymd(paste0(year, "-6-30")),
    quarter == 3 ~ ymd(paste0(year, "-9-30")),
    quarter == 4 ~ ymd(paste0(year, "-12-31"))
  )) %>%
  filter(year >= 2015) %>%
  mutate(
    time = format(as.POSIXct(crime_date_time), format =  "%I:%M:%S %p")
  )  %>%
  mutate(during_or_just_after_school = case_when(
    hour >= 7 & hour <= 18 ~ TRUE,
    TRUE ~ FALSE
  ))

shooting_victims <- victims %>%
  #filtering for shootings and homicides where a firearm was used
  filter(description %in% c("shooting", "homicide")) %>%
  filter(weapon == "firearm")

```
## MUTATE TO UNIQUE CRIME EVENTS

```{r}

shooting_crimes <- shooting_victims %>%
  #grouping by distinct instance characteristics to counts crime incidents instead of victions
  group_by(crime_date_time, location, district, neighborhood, premise, year, month, date, time_range, quarter, latitude, longitude) %>%
  #summarising characteristics of crimes victims
  summarise(victims = n(),
            child_victims = sum(age_range == "child"),
            child_homicides = sum(description == "homicide" & age_range == "child"),
            high_school_teen_homicides = sum(description == "homicide" & age_range == "juvenile"),
            high_school_teen_victims = sum(age_range == "high school teen"),
            homicides = sum(description == "homicide"),
            juvenile_homicides = sum(description == "homicide" & age_range == "juvenile"),
            average_age = mean(age, na.rm = TRUE),
            description = paste0(unique(description), collapse = ", "),
            age_of_victims = paste0(age, collapse = ", "),
            black_victims = sum(race == "black_or_african_american"),
            white_victims = sum(race == "white"),
            asian_victims = sum(race == "asian")
            )

years <- 2015:2023

```

## IMPORT SCHOOL DATABASE AND IDENTIFY PARCELS

```{r}

### neighborhood shapefiles for map later
neighborhoods <- st_read("data/shapefiles/neighborhood/neighborhood.shp") %>%
  st_transform(crs = 4326) %>%
  clean_names() #%>%
#  rename( 
#    neighborhood = label)


school_list <- read_csv("data/schoollist.csv") %>% 
  clean_names() %>%
  lower_df() %>%
  ## manually editing one longitude
  mutate(
    address_longitude = case_when(address_longitude == "--76.559637" ~	"-76.559637",
                                  TRUE ~ as.character(address_longitude))
  ) 

## high schools + making spatial object
spatial_school_list <- school_list %>%
  #filter(str_detect(official_state_grade_band, "high")) %>%
  st_as_sf(coords = c("address_longitude", "address_latitude"),
                        crs = 4326) %>%
  ## putting filter up here for schools we don't care about
  filter(
    school_number != 875,
    school_number != 877,
    school_number != 878,
    school_number != 884,
    school_number != 897
  )

  
parcels <- st_read("data/shapefiles/Real_Property_Information/Real_Property_Information.shp") %>%
  st_transform(crs = 4326) %>%
  clean_names() %>%
  lower_df()

### this makes join work for some reason
sf_use_s2(FALSE)

### check for if school_geoms are in parcel. 
schools_with_parcels <- parcels %>%
    st_join(spatial_school_list) 

school_parcels <- schools_with_parcels %>%
  select(school_number, school_name, address, fulladdr) %>%
  drop_na(school_name) %>%
  distinct(school_number, .keep_all = TRUE)

# check that shapes look okay
map <- mapview(school_parcels)+ mapview(spatial_school_list, cex = 2, color = "red")
map

### removing some schools that we're not considering because students don't leave the property (school is in a prison) or several schools are in one building, or the schools are virtual.
# p-tech at carver vocational-technical high, p-tech at paul laurence dunbar high, p-tech at new era academy, port vlp at digital harbor, eager street academy

final_parcels <- school_parcels %>%
  filter(
    school_number != 875,
    school_number != 877,
    school_number != 878,
    school_number != 884,
    school_number != 897
  )

## some duplicate parcels here, so distincting based on parcel address duplicates columnn (full_addr)
final_parcels <- final_parcels %>%
  group_by(fulladdr) %>%
  mutate(
    schools_at_parcel = paste0(str_to_title(school_name), collapse = ", ")
  ) %>%
  distinct(fulladdr, .keep_all = TRUE)
  
#sanity check 
mapview(final_parcels) + mapview(spatial_school_list, cex = 2, color = "red")
# 138 unique parcels!

```


"High-school-age teens in Baltimore continue to be shot in record numbers this year, even as overall, nonfatal shootings and homicides are down, according to an analysis by The Baltimore Banner."

```{r}
  
for (nume in seq(1:length(years))){
  
  input_year <- years[nume]
  
  temp <- shooting_victims %>% 
    filter(year == input_year,
           date <= mdy(paste0("2/18/", input_year))) %>%
    mutate(temp_high_school_age = case_when(
      age_range %notin% c("high school teen", "child") ~ "19 and older",
      TRUE ~ age_range
    )) %>%
    group_by(temp_high_school_age) %>%
    summarise(!!paste0("x", input_year) := n()) 
  
  if (nrow(temp %>% filter(temp_high_school_age == "child")) == 0){
  
  temp <- temp %>%
    bind_rows(data.frame(temp_high_school_age = "child", count = 0) %>%
                rename(!!paste0("x", input_year) := count)) %>%
    arrange(temp_high_school_age)
  
  }
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "high_school_teen_share", temp = (temp[2][[1]][3]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "child_share", temp = (temp[2][[1]][2]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
    
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "19_and_over_share", temp = (temp[2][[1]][1]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
  
  if (nume == 1){
    
    output <- temp
    
  } else {
    output <- output %>%
      full_join(temp)
  }
}

output[is.na(output)] <- 0

raw_high_school_counts_by_year_through_feb_18 <- output %>%
  round_df(2)


raw_high_school_counts_by_year_through_feb_18

```

"Last year was a record year for children being shot in Baltimore, a disturbing trend first reported by the Banner in September. That year ended with 84 juvenile shooting victims. But since then, the situation has gotten worse."

```{r}

shooting_victims %>%
  filter(age <= 17) %>%
  group_by(year) %>%
  count()

```

"January had the most high school-aged teenage victims of gun violence of any month since 2015, the first year with reliable publicly available data."

```{r}

shooting_crimes %>%
  filter(high_school_teen_victims > 0) %>%
  group_by(month, year) %>%
  summarise(high_school_teen_victims = sum(high_school_teen_victims),
            high_school_teen_shootings = n()) %>%
  arrange(desc(high_school_teen_victims))

```
"The two most violent weeks since 2015 occurred this year: the first week of January and the second week of February."

```{r}

shooting_crimes %>%
  mutate(week = week(date)) %>%
  filter(high_school_teen_victims > 0) %>%
  group_by(week, year) %>%
  summarise(high_school_teen_victims = sum(high_school_teen_victims),
            high_school_teen_shootings = n()) %>%
  arrange(desc(high_school_teen_victims))

```

"Last year, the shootings were concentrated in the city’s Eastern District."

```{r}

shooting_crimes %>%
  filter(high_school_teen_victims > 0,
         year == 2022) %>%
  group_by(district) %>%
  summarise(high_school_teen_victims = sum(high_school_teen_victims),
            high_school_teen_shootings = n()) %>%
  arrange(desc(high_school_teen_victims))

```

"So far this year, the violence has spread to the Northeastern District and taken hold in the Southwestern District, which was primarily driven by a mass shooting across the street from Edmondson-Westside High School that killed one teenager and injured four."

```{r}

shooting_crimes %>%
  filter(high_school_teen_victims > 0,
         year == 2023) %>%
  group_by(district) %>%
  summarise(high_school_teen_victims = sum(high_school_teen_victims),
            high_school_teen_shootings = n()) %>%
  arrange(desc(high_school_teen_victims))

```

#### NEAR SCHOOLS ANALYSIS

```{r}

shootings <- victims %>%
  filter(description %in% c("shooting", "homicide"),
         weapon == "firearm") %>%
  #mutating database of victims to database of distinct crimes
  group_by(crime_date_time, location, district, neighborhood, race, weapon, premise, year, month, date, time_range, hour, during_or_just_after_school, quarter, longitude, latitude) %>%
  summarise(victims = n(),
            child_victims = sum(age_range == "child"),
            child_homicides = sum(description == "homicide" & age_range == "child"),
            high_school_teen_homicides = sum(description == "homicide" & age_range == "high school teen"),
            high_school_teen_victims = sum(age_range == "high school teen"),
            homicides = sum(description == "homicide"),
            average_age = mean(age, na.rm = TRUE),
            description = paste0(unique(description), collapse = ", "),
            age_of_victims = paste0(age, collapse = ", "),
            black_victims = sum(race == "black_or_african_american"),
            white_victims = sum(race == "white"),
            asian_victims = sum(race == "asian")
            ) %>%
  filter(longitude != "",
         latitude != "") %>%
  ungroup() %>%
  mutate(academic_year = case_when(
      date >= '2015-09-01' & date <= '2016-06-15' ~ "2015 - 2016",
      date >= '2016-09-01' & date <= '2017-06-15' ~ "2016 - 2017",
      date >= '2017-09-01' & date <= '2018-06-15' ~ "2017 - 2018",
      date >= '2018-09-01' & date <= '2019-06-15' ~ "2018 - 2019",
      date >= '2019-09-01' & date <= '2020-06-15' ~ "2019 - 2020",
      date >= '2020-09-01' & date <= '2021-06-15' ~ "2020 - 2021",
      date >= '2021-09-01' & date <= '2022-06-15' ~ "2021 - 2022",
      date >= '2022-09-01' & date <= '2023-06-15' ~ "2022 - 2023"
    )) %>%
  mutate(
    time = format(as.POSIXct(crime_date_time), format =  "%I:%M:%S %p")
  )  %>%
  st_as_sf(coords = c("longitude", "latitude"),
                 crs = 4326,#4269, 
                 remove = FALSE)

  school_shape <- final_parcels

  school_shape$school_number <- rownames(school_shape)

  blocks <- 2
  
  distance <- blocks*100+50  

near_school <- as.data.frame(st_is_within_distance(shootings, school_shape, dist = distance, sparse = FALSE)) %>%
  clean_names() %>%
  mutate(within = case_when(
     v1 == TRUE ~ TRUE,
     v2 == TRUE ~ TRUE,
     v3 == TRUE ~ TRUE,
     v4 == TRUE ~ TRUE,
     v5 == TRUE ~ TRUE,
     v6 == TRUE ~ TRUE,
     v7 == TRUE ~ TRUE,
     v8 == TRUE ~ TRUE,
     v9 == TRUE ~ TRUE,
     v10 == TRUE ~ TRUE,
     v11 == TRUE ~ TRUE,
     v12 == TRUE ~ TRUE,
     v13 == TRUE ~ TRUE,
     v14 == TRUE ~ TRUE,
     v15 == TRUE ~ TRUE,
     v16 == TRUE ~ TRUE,
     v17 == TRUE ~ TRUE,
     v18 == TRUE ~ TRUE,
     v19 == TRUE ~ TRUE,
     v20 == TRUE ~ TRUE,
     v21 == TRUE ~ TRUE,
     v22 == TRUE ~ TRUE,
     v23 == TRUE ~ TRUE,
     v24 == TRUE ~ TRUE,
     v25 == TRUE ~ TRUE,
     v26 == TRUE ~ TRUE,
     v27 == TRUE ~ TRUE,
     v28 == TRUE ~ TRUE,
     v29 == TRUE ~ TRUE,
     v30 == TRUE ~ TRUE,
     v31 == TRUE ~ TRUE,
     v32 == TRUE ~ TRUE,
     v33 == TRUE ~ TRUE,
     v34 == TRUE ~ TRUE,
     v35 == TRUE ~ TRUE,
     v36 == TRUE ~ TRUE,
     v37 == TRUE ~ TRUE,
     v38 == TRUE ~ TRUE,
     v39 == TRUE ~ TRUE,
     v40 == TRUE ~ TRUE,
     v41 == TRUE ~ TRUE,
     v42 == TRUE ~ TRUE,
     v43 == TRUE ~ TRUE,
     v44 == TRUE ~ TRUE,
     v45 == TRUE ~ TRUE,
     v46 == TRUE ~ TRUE,
     v47 == TRUE ~ TRUE,
     v48 == TRUE ~ TRUE,
     v49 == TRUE ~ TRUE,
     v50 == TRUE ~ TRUE,
     v51 == TRUE ~ TRUE,
     v52 == TRUE ~ TRUE,
     v53 == TRUE ~ TRUE,
     v54 == TRUE ~ TRUE,
     v55 == TRUE ~ TRUE,
     v56 == TRUE ~ TRUE,
     v57 == TRUE ~ TRUE,
     v58 == TRUE ~ TRUE,
     v59 == TRUE ~ TRUE,
     v60 == TRUE ~ TRUE,
     v61 == TRUE ~ TRUE,
     v62 == TRUE ~ TRUE,
     v63 == TRUE ~ TRUE,
     v64 == TRUE ~ TRUE,
     v65 == TRUE ~ TRUE,
     v66 == TRUE ~ TRUE,
     v67 == TRUE ~ TRUE,
     v68 == TRUE ~ TRUE,
     v69 == TRUE ~ TRUE,
     v70 == TRUE ~ TRUE,
     v71 == TRUE ~ TRUE,
     v72 == TRUE ~ TRUE,
     v73 == TRUE ~ TRUE,
     v74 == TRUE ~ TRUE,
     v75 == TRUE ~ TRUE,
     v76 == TRUE ~ TRUE,
     v77 == TRUE ~ TRUE,
     v78 == TRUE ~ TRUE,
     v79 == TRUE ~ TRUE,
     v80 == TRUE ~ TRUE,
     v89 == TRUE ~ TRUE,
     v90 == TRUE ~ TRUE,
     v91 == TRUE ~ TRUE,
     v92 == TRUE ~ TRUE,
     v93 == TRUE ~ TRUE,
     v94 == TRUE ~ TRUE,
     v95 == TRUE ~ TRUE,
     v96 == TRUE ~ TRUE,
     v97 == TRUE ~ TRUE,
     v98 == TRUE ~ TRUE,
     v99 == TRUE ~ TRUE,
     v100 == TRUE ~ TRUE,
     v101 == TRUE ~ TRUE,
     v102 == TRUE ~ TRUE,
     v103 == TRUE ~ TRUE,
     v104 == TRUE ~ TRUE,
     v105 == TRUE ~ TRUE,
     v106 == TRUE ~ TRUE,
     v107 == TRUE ~ TRUE,
     v108 == TRUE ~ TRUE,
     v109 == TRUE ~ TRUE,
     v110 == TRUE ~ TRUE,
     v111 == TRUE ~ TRUE,
     v112 == TRUE ~ TRUE,
     v113 == TRUE ~ TRUE,
     v114 == TRUE ~ TRUE,
     v115 == TRUE ~ TRUE,
     v116 == TRUE ~ TRUE,
     v117 == TRUE ~ TRUE,
     v118 == TRUE ~ TRUE,
     v119 == TRUE ~ TRUE,
     v110 == TRUE ~ TRUE,
     v111 == TRUE ~ TRUE,
     v112 == TRUE ~ TRUE,
     v113 == TRUE ~ TRUE,
     v114 == TRUE ~ TRUE,
     v115 == TRUE ~ TRUE,
     v116 == TRUE ~ TRUE,
     v117 == TRUE ~ TRUE,
     v118 == TRUE ~ TRUE,
     v119 == TRUE ~ TRUE,
     v120 == TRUE ~ TRUE,
     v121 == TRUE ~ TRUE,
     v122 == TRUE ~ TRUE,
     v123 == TRUE ~ TRUE,
     v124 == TRUE ~ TRUE,
     v125 == TRUE ~ TRUE,
     v126 == TRUE ~ TRUE,
     v127 == TRUE ~ TRUE,
     v128 == TRUE ~ TRUE,
     v129 == TRUE ~ TRUE,
     v130 == TRUE ~ TRUE,
     v131 == TRUE ~ TRUE,
     v132 == TRUE ~ TRUE,
     v133 == TRUE ~ TRUE,
     v134 == TRUE ~ TRUE,
     v135 == TRUE ~ TRUE,
     v136 == TRUE ~ TRUE,
     v137 == TRUE ~ TRUE,
     v138 == TRUE ~ TRUE,
     TRUE ~ FALSE
  )) %>%
  #select(within) %>%
  mutate(school_number = case_when(
     v1 == TRUE ~ 1,
     v2 == TRUE ~ 2,
     v3 == TRUE ~ 3,
     v4 == TRUE ~ 4,
     v5 == TRUE ~ 5,
     v6 == TRUE ~ 6,
     v7 == TRUE ~ 7,
     v8 == TRUE ~ 8,
     v9 == TRUE ~ 9,
     v10 == TRUE ~ 10,
     v11 == TRUE ~ 11,
     v12 == TRUE ~ 12,
     v13 == TRUE ~ 13,
     v14 == TRUE ~ 14,
     v15 == TRUE ~ 15,
     v16 == TRUE ~ 16,
     v17 == TRUE ~ 17,
     v18 == TRUE ~ 18,
     v19 == TRUE ~ 19,
     v20 == TRUE ~ 20,
     v21 == TRUE ~ 21,
     v22 == TRUE ~ 22,
     v23 == TRUE ~ 23,
     v24 == TRUE ~ 24,
     v25 == TRUE ~ 25,
     v26 == TRUE ~ 26,
     v27 == TRUE ~ 27,
     v28 == TRUE ~ 28,
     v29 == TRUE ~ 29,
     v30 == TRUE ~ 30,
     v31 == TRUE ~ 31,
     v32 == TRUE ~ 32,
     v33 == TRUE ~ 33,
     v34 == TRUE ~ 34,
     v35 == TRUE ~ 35,
     v36 == TRUE ~ 36,
     v37 == TRUE ~ 37,
     v38 == TRUE ~ 38,
     v39 == TRUE ~ 39,
     v40 == TRUE ~ 40,
     v41 == TRUE ~ 41,
     v42 == TRUE ~ 42,
     v43 == TRUE ~ 43,
     v44 == TRUE ~ 44,
     v45 == TRUE ~ 45,
     v46 == TRUE ~ 46,
     v47 == TRUE ~ 47,
     v48 == TRUE ~ 48,
     v49 == TRUE ~ 49,
     v50 == TRUE ~ 50,
     v51 == TRUE ~ 51,
     v52 == TRUE ~ 52,
     v53 == TRUE ~ 53,
     v54 == TRUE ~ 54,
     v55 == TRUE ~ 55,
     v56 == TRUE ~ 56,
     v57 == TRUE ~ 57,
     v58 == TRUE ~ 58,
     v59 == TRUE ~ 59,
     v60 == TRUE ~ 60,
     v61 == TRUE ~ 61,
     v62 == TRUE ~ 62,
     v63 == TRUE ~ 63,
     v64 == TRUE ~ 64,
     v65 == TRUE ~ 65,
     v66 == TRUE ~ 66,
     v67 == TRUE ~ 67,
     v68 == TRUE ~ 68,
     v69 == TRUE ~ 69,
     v70 == TRUE ~ 70,
     v71 == TRUE ~ 71,
     v72 == TRUE ~ 72,
     v73 == TRUE ~ 73,
     v74 == TRUE ~ 74,
     v75 == TRUE ~ 75,
     v76 == TRUE ~ 76,
     v77 == TRUE ~ 77,
     v78 == TRUE ~ 78,
     v79 == TRUE ~ 79,
     v80 == TRUE ~ 80,
     v89 == TRUE ~ 89,
     v90 == TRUE ~ 90,
     v91 == TRUE ~ 91,
     v92 == TRUE ~ 92,
     v93 == TRUE ~ 93,
     v94 == TRUE ~ 94,
     v95 == TRUE ~ 95,
     v96 == TRUE ~ 96,
     v97 == TRUE ~ 97,
     v98 == TRUE ~ 98,
     v99 == TRUE ~ 99,
     v100 == TRUE ~ 100,
     v101 == TRUE ~ 101,
     v102 == TRUE ~ 102,
     v103 == TRUE ~ 103,
     v104 == TRUE ~ 104,
     v105 == TRUE ~ 105,
     v106 == TRUE ~ 106,
     v107 == TRUE ~ 107,
     v108 == TRUE ~ 108,
     v109 == TRUE ~ 109,
     v110 == TRUE ~ 110,
     v111 == TRUE ~ 111,
     v112 == TRUE ~ 112,
     v113 == TRUE ~ 113,
     v114 == TRUE ~ 114,
     v115 == TRUE ~ 115,
     v116 == TRUE ~ 116,
     v117 == TRUE ~ 117,
     v118 == TRUE ~ 118,
     v119 == TRUE ~ 119,
     v110 == TRUE ~ 110,
     v111 == TRUE ~ 111,
     v112 == TRUE ~ 112,
     v113 == TRUE ~ 113,
     v114 == TRUE ~ 114,
     v115 == TRUE ~ 115,
     v116 == TRUE ~ 116,
     v117 == TRUE ~ 117,
     v118 == TRUE ~ 118,
     v119 == TRUE ~ 119,
     v120 == TRUE ~ 120,
     v121 == TRUE ~ 121,
     v122 == TRUE ~ 122,
     v123 == TRUE ~ 123,
     v124 == TRUE ~ 124,
     v125 == TRUE ~ 125,
     v126 == TRUE ~ 126,
     v127 == TRUE ~ 127,
     v128 == TRUE ~ 128,
     v129 == TRUE ~ 129,
     v130 == TRUE ~ 130,
     v131 == TRUE ~ 131,
     v132 == TRUE ~ 132,
     v133 == TRUE ~ 133,
     v134 == TRUE ~ 134,
     v135 == TRUE ~ 135,
     v136 == TRUE ~ 136,
     v137 == TRUE ~ 137,
     v138 == TRUE ~ 138,
     TRUE ~ FALSE
  )) %>%
  select(within, school_number)


near_school <- shootings %>%
  cbind(near_school$within, near_school$school_number) %>%
  clean_names() %>%
  mutate(near_school_school_number = as.character(near_school_school_number)) %>%
  filter(near_school_within == TRUE) %>%
  filter(during_or_just_after_school == "TRUE") %>%
  #filter(high_school_teen_victims > 0) %>%
  #filter(description == "shooting" | 
  #       description == "homicide") %>%
  filter(weapon == "firearm") %>%
  as.data.frame() %>%
  left_join(school_shape %>%
              select(schools_at_parcel, school_number),
            by = c("near_school_school_number" = "school_number"))
  
  near_school_by_year <- near_school %>%
    as.data.frame() %>%
    #mutate(year_month = mdy(paste0(month, "-1-", year))) %>%
    group_by(year) %>%
    summarise(shootings = n(),
              high_school_teen_victims = sum(high_school_teen_victims),
              total_victims = sum(victims))
  
  near_school_academic_year <- near_school %>%
    as.data.frame() %>%
    #mutate(year_month = mdy(paste0(month, "-1-", year))) %>%
    group_by(academic_year) %>%
    summarise(shootings = n(),
              high_school_teen_victims = sum(high_school_teen_victims),
              total_victims = sum(victims))
  
  plot <- ggplot(near_school_by_year, aes(x=year, y=shootings)) +
    geom_line() +
    ggtitle(paste0("shootings within ", blocks, " block(s) of 31 high schools during school hours"))
  
  near_school_by_hour <- near_school %>%
    mutate(crime_date_time = ymd_hms(crime_date_time))  %>%
    mutate(year = year(crime_date_time),
         month = month(crime_date_time),
         hour = hour(crime_date_time)) %>%
    group_by(hour) %>%
    summarize(
      count = n()
    )

### writing out two blocks for viz 

clean_near_school <- near_school %>%
  mutate(
    location = str_to_title(location),
    neighborhood = str_to_title(neighborhood),
    premise = str_to_title(premise),
    district = str_to_title(district),
    time_range = str_to_title(time_range)
  ) %>%
  filter(!is.na(academic_year))

clean_parcels <- final_parcels %>%
  mutate(
    school_name = str_to_title(school_name),
    fulladdr = str_to_title(fulladdr),
  )

```


"Twenty-three school-age children and young adults, ages 13 to 18, have been shot near a Baltimore City school during the school day this academic year."

```{r}

near_school_academic_year %>%
  filter(academic_year == "2022 - 2023") %>%
  select(academic_year, high_school_teen_victims)

```

"Winter has been especially violent. More than half of all of the victims were shot since the winter break.

```{r}

near_school %>%
  filter(academic_year == "2022 - 2023") %>%
  group_by(year) %>%
  summarise(high_school_teen_victims = sum(high_school_teen_victims))

```

"Nearly 1 in 4 juvenile gunshot victims were shot near a school this academic year, and 18-year-olds were shot near schools at similar rates as 17-year-olds."

```{r}

#september 1 to june 15
shooting_victims_by_school_year <- shooting_victims %>%
  mutate(school_year = case_when(
    date > mdy("6/15/2015") & date <= mdy("6/15/2016") ~ "2015/2016",
    date > mdy("6/15/2016") & date <= mdy("6/15/2017") ~ "2016/2017",
    date > mdy("6/15/2017") & date <= mdy("6/15/2018") ~ "2017/2018",
    date > mdy("6/15/2018") & date <= mdy("6/15/2019") ~ "2018/2019",    
    date > mdy("6/15/2019") & date <= mdy("6/15/2020") ~ "2019/2020",
    date > mdy("6/15/2020") & date <= mdy("6/15/2021") ~ "2020/2021",
    date > mdy("6/15/2021") & date <= mdy("6/15/2022") ~ "2021/2022",
    date > mdy("6/15/2022") & date <= mdy("6/15/2023") ~ "2022/2023",
    TRUE ~ "error"
  ))

school_years <- c("2015/2016", "2016/2017", "2017/2018", "2018/2019", "2019/2020", "2020/2021", "2021/2022", "2022/2023")

for (nume in seq(1:length(school_years))){
  
  input_year <- school_years[nume]
  
  end_year <- str_split(input_year, "/")[[1]][2]
  
  temp <- shooting_victims_by_school_year %>% 
    filter(school_year == input_year,
           date <= mdy(paste0("2/18/", end_year))) %>%
    mutate(temp_high_school_age = case_when(
      age_range %notin% c("high school teen", "child") ~ "19 and older",
      TRUE ~ age_range
    )) %>%
    group_by(temp_high_school_age) %>%
    summarise(!!paste0("x", input_year) := n()) 
  
  if (nrow(temp %>% filter(temp_high_school_age == "child")) == 0){
  
  temp <- temp %>%
    bind_rows(data.frame(temp_high_school_age = "child", count = 0) %>%
                rename(!!paste0("x", input_year) := count)) %>%
    arrange(temp_high_school_age)
  
  }
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "high_school_teen_share", temp = (temp[2][[1]][3]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "child_share", temp = (temp[2][[1]][2]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
    
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "19_and_over_share", temp = (temp[2][[1]][1]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
    
  if (nume == 1){
    
    output <- temp
    
  } else {
    output <- output %>%
      full_join(temp)
  }
}

output[is.na(output)] <- 0


high_school_age_victims_by_school_year_thru_feb_18 <- output %>%
  round_df(2)

high_school_age_victims_by_school_year_thru_feb_18 %>%
  clean_names() %>%
  select(temp_high_school_age, x2022_2023) %>%
  filter(temp_high_school_age == "high school teen")

#write_csv(high_school_age_victims_by_school_year_thru_feb_18 %>%
            #filter(boolean_high_school != "Total"), "data/output/high_school_age_victims_by_school_year_thru_feb_18.csv")

```

```{r}
#high school teens shot near schools during school year, total high school teens shot since end of last school year
(23/89)*100

```

"Very few children 12 and under are shot in Baltimore, and the rate is actually down significantly since 2015."

```{r}

shooting_victims %>%
  filter(age_range == "child") %>%
  group_by(year) %>%
  count()

```


### VIZ

# SHARE OVER TIME VIZ

##################################################################
### HIGH SCHOOL SHOOTING VICTIMS CALENDAR YEAR THROUGH FEB. 18 ###
##################################################################

```{r}
  
for (nume in seq(1:length(years))){
  
  input_year <- years[nume]
  
  temp <- shooting_victims %>% 
    filter(year == input_year,
           date <= mdy(paste0("2/18/", input_year))) %>%
    mutate(temp_high_school_age = case_when(
      age_range %notin% c("high school teen", "child") ~ "19 and older",
      TRUE ~ age_range
    )) %>%
    group_by(temp_high_school_age) %>%
    summarise(!!paste0("x", input_year) := n()) 
  
  if (nrow(temp %>% filter(temp_high_school_age == "child")) == 0){
  
  temp <- temp %>%
    bind_rows(data.frame(temp_high_school_age = "child", count = 0) %>%
                rename(!!paste0("x", input_year) := count)) %>%
    arrange(temp_high_school_age)
  
  }
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "high_school_teen_share", temp = (temp[2][[1]][3]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "child_share", temp = (temp[2][[1]][2]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
    
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "19_and_over_share", temp = (temp[2][[1]][1]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
  
  if (nume == 1){
    
    output <- temp
    
  } else {
    output <- output %>%
      full_join(temp)
  }
}

output[is.na(output)] <- 0

raw_high_school_counts_by_year_through_feb_18 <- output %>%
  round_df(2)


raw_high_school_counts_by_year_through_feb_18

```

#######################################################
### SHOOTING VICTIMS THROUGH FEB. 15 BY SCHOOL YEAR ###
#######################################################

```{r}

#september 1 to june 15
shooting_victims_by_school_year <- shooting_victims %>%
  mutate(school_year = case_when(
    date > mdy("6/15/2015") & date <= mdy("6/15/2016") ~ "2015/2016",
    date > mdy("6/15/2016") & date <= mdy("6/15/2017") ~ "2016/2017",
    date > mdy("6/15/2017") & date <= mdy("6/15/2018") ~ "2017/2018",
    date > mdy("6/15/2018") & date <= mdy("6/15/2019") ~ "2018/2019",    
    date > mdy("6/15/2019") & date <= mdy("6/15/2020") ~ "2019/2020",
    date > mdy("6/15/2020") & date <= mdy("6/15/2021") ~ "2020/2021",
    date > mdy("6/15/2021") & date <= mdy("6/15/2022") ~ "2021/2022",
    date > mdy("6/15/2022") & date <= mdy("6/15/2023") ~ "2022/2023",
    TRUE ~ "error"
  ))

school_years <- c("2015/2016", "2016/2017", "2017/2018", "2018/2019", "2019/2020", "2020/2021", "2021/2022", "2022/2023")

for (nume in seq(1:length(school_years))){
  
  input_year <- school_years[nume]
  
  end_year <- str_split(input_year, "/")[[1]][2]
  
  temp <- shooting_victims_by_school_year %>% 
    filter(school_year == input_year,
           date <= mdy(paste0("2/18/", end_year))) %>%
    mutate(temp_high_school_age = case_when(
      age_range %notin% c("high school teen", "child") ~ "19 and older",
      TRUE ~ age_range
    )) %>%
    group_by(temp_high_school_age) %>%
    summarise(!!paste0("x", input_year) := n()) 
  
  if (nrow(temp %>% filter(temp_high_school_age == "child")) == 0){
  
  temp <- temp %>%
    bind_rows(data.frame(temp_high_school_age = "child", count = 0) %>%
                rename(!!paste0("x", input_year) := count)) %>%
    arrange(temp_high_school_age)
  
  }
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "high_school_teen_share", temp = (temp[2][[1]][3]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "child_share", temp = (temp[2][[1]][2]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
    
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "19_and_over_share", temp = (temp[2][[1]][1]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
    
  if (nume == 1){
    
    output <- temp
    
  } else {
    output <- output %>%
      full_join(temp)
  }
}

output[is.na(output)] <- 0


high_school_age_victims_by_school_year_thru_feb_18 <- output %>%
  round_df(2)

high_school_age_victims_by_school_year_thru_feb_18

```

#################################################
### SHOOTING VICTIMS BY LAST 12 MONTH PERIODS ###
#################################################

```{r}

#september 1 to june 15
shooting_victims_by_12_month_periods <- shooting_victims %>%
  mutate(x12_month_period = case_when(
    date > mdy("2/18/2015") & date <= mdy("2/18/2016") ~ "2015/2016",
    date > mdy("2/18/2016") & date <= mdy("2/18/2017") ~ "2016/2017",
    date > mdy("2/18/2017") & date <= mdy("2/18/2018") ~ "2017/2018",
    date > mdy("2/18/2018") & date <= mdy("2/18/2019") ~ "2018/2019",    
    date > mdy("2/18/2019") & date <= mdy("2/18/2020") ~ "2019/2020",
    date > mdy("2/18/2020") & date <= mdy("2/18/2021") ~ "2020/2021",
    date > mdy("2/18/2021") & date <= mdy("2/18/2022") ~ "2021/2022",
    date > mdy("2/18/2022") & date <= mdy("2/18/2023") ~ "2022/2023",
    TRUE ~ "error"
  ))

school_years <- c("2015/2016", "2016/2017", "2017/2018", "2018/2019", "2019/2020", "2020/2021", "2021/2022", "2022/2023")

for (nume in seq(1:length(school_years))){
  
  input_year <- school_years[nume]
  
  temp <- shooting_victims_by_12_month_periods %>% 
    filter(x12_month_period == input_year) %>%
    mutate(temp_high_school_age = case_when(
      age_range %notin% c("high school teen", "child") ~ "19 and older",
      TRUE ~ age_range
    )) %>%
    group_by(temp_high_school_age) %>%
    summarise(!!paste0("x", input_year) := n()) 
  
  if (nrow(temp %>% filter(temp_high_school_age == "child")) == 0){
  
  temp <- temp %>%
    bind_rows(data.frame(temp_high_school_age = "child", count = 0) %>%
                rename(!!paste0("x", input_year) := count)) %>%
    arrange(temp_high_school_age)
  
  }
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "high_school_teen_share", temp = (temp[2][[1]][3]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "child_share", temp = (temp[2][[1]][2]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
    
  temp <- temp%>%
    bind_rows(data.frame(temp_high_school_age = "19_and_over_share", temp = (temp[2][[1]][1]/(temp[2][[1]][1]+temp[2][[1]][2]+temp[2][[1]][3]))*100) %>%
                rename(!!paste0("x", input_year) := temp))
  
    
  if (nume == 1){
    
    output <- temp
    
  } else {
    output <- output %>%
      full_join(temp)
  }
}

output[is.na(output)] <- 0

high_school_age_victims_in_last_12_month_periods <- output %>%
  round_df(2)

high_school_age_victims_in_last_12_month_periods

```

## BUILDING OUTPUT FOR VIZ

```{r}

temp <- raw_high_school_counts_by_year_through_feb_18 %>%
  pivot_longer(names_to = "year", cols = c(x2015, x2016, x2017, x2018, x2019, x2020, x2021, x2022, x2023)) %>%
  pivot_wider(values_from = value, names_from = temp_high_school_age) %>%
  mutate(data_type = "Jan. 1 through Feb. 18",
         year = gsub("x", "", year)) 


temp2 <- high_school_age_victims_by_school_year_thru_feb_18 %>%
  clean_names() %>%
  pivot_longer(names_to = "year", cols = c(x2015_2016, x2016_2017, x2017_2018, x2018_2019, x2019_2020, x2020_2021, x2021_2022, x2022_2023)) %>%
  pivot_wider(values_from = value, names_from = temp_high_school_age) %>%
  mutate(data_type = "School year through Feb. 15",
         year = gsub("x", "", year),
         year = gsub("_", "/", year)) 

temp3 <- high_school_age_victims_in_last_12_month_periods %>%
  clean_names() %>%
  pivot_longer(names_to = "year", cols = c(x2015_2016, x2016_2017, x2017_2018, x2018_2019, x2019_2020, x2020_2021, x2021_2022, x2022_2023)) %>%
  pivot_wider(values_from = value, names_from = temp_high_school_age) %>%
  mutate(data_type = "12-month period ending Feb. 18",
         year = gsub("x", "", year),
         year = gsub("_", "/", year)) 

```

#######################################
### SHOOTINGS BY YEAR FOR STORY MAP ###
#######################################

```{r}

for (input_year in (2015:2023)){
  temp <- shooting_crimes %>%
           filter(high_school_teen_victims > 0,
                  year == input_year) %>%
  mutate(location = str_to_title(location),
         district = str_to_title(district),
         neighborhood = str_to_title(neighborhood),
         premise = str_to_title(premise),
         time_range = str_to_title(time_range)
         ) %>%
    filter(!is.na(latitude),
           !is.na(neighborhood))
  
  #temp <- st_as_sf(temp,
  #                 coords = c("longitude", "latitude"),
  #                 crs = 4326)
  
  #write_csv(temp, paste0("data/output/high-school-aged-victims-", input_year,".csv"))
  
}

```
